---
title: "TP1"
author: "COURIVAUD - KOUEK"
date: "24 novembre 2016"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Data preparation

```{r }
FluTrain <- read.csv("FluTrain.csv", header=FALSE)
colnames(FluTrain) = c("Weeks", "ILI", "Queries")
```

```{r message = FALSE}
library(tidyr)
library(hydroGOF)
```

## Data Date Formating
```{r }
FluTrain = separate(data = FluTrain, col = Weeks, into = c("date1", "date2"), sep = " - " )
FluTrain$date1 = strptime(FluTrain$date1,format="%Y-%m-%d")
FluTrain$date2 = strptime(FluTrain$date2,format="%Y-%m-%d")
FluTrain$year = FluTrain$date1$year + 1900
FluTrain$week = as.numeric( format(FluTrain$date1+3, "%U"))
FluTrain$day = FluTrain$date1$nday
FluTrain$month = FluTrain$date1$mon+1
```
### Nombre de cas par an


```{r }
plot(FluTrain$ILI,FluTrain$Queries)
```


### Cycle of Illness over years 
```{r }
plot(FluTrain$ILI/mean(FluTrain$ILI), col="blue")
lines(FluTrain$Queries/mean(FluTrain$Queries), col="red")
```


# 2
## 2.1.1

The FluTrain dataset period is in 2004-2011.

### Question 1)
```{r }
FluTrain[FluTrain$ILI==max(FluTrain$ILI),]

paste("The week is ",FluTrain[FluTrain$ILI==max(FluTrain$ILI),]$week)
paste("The day is ",FluTrain[FluTrain$ILI==max(FluTrain$ILI),]$date1$mday)
paste("The month is ",FluTrain[FluTrain$ILI==max(FluTrain$ILI),]$date1$mon+1)
```

### Question 2)
```{r }
FluTrain[FluTrain$Queries==max(FluTrain$Queries),]

paste("The week is ",FluTrain[FluTrain$Queries==max(FluTrain$Queries),]$week)
paste("The day is ",FluTrain[FluTrain$Queries==max(FluTrain$Queries),]$date1$mday)
paste("The month is ",FluTrain[FluTrain$Queries==max(FluTrain$Queries),]$date1$mon+1)
```

### 2.1.2
```{r }
histoILI = hist(FluTrain$ILI)
```

Most of the ILI values are small, with a relatively small number of much larger values (in statistics, this sort of data is called "skew right").

### 2.2.3
```{r }
plot((log(FluTrain$ILI)~(FluTrain$Queries)))
```

There is a positive, linear relationship between log(ILI) and Queries.

### 2.2.4
log(ILI) = β 0 + β 1 × Queries, where the slope β 1 is positive.

### 2.2.5
```{r }
FluTrend1 = lm(log(ILI) ~ Queries, data=FluTrain)
summary(FluTrend1)

mseTrend1.1 = sum(FluTrend1$residuals**2)/length(FluTrend1$residuals)
mseTrend1.2 = sum(((predict(FluTrend1, FluTrain)-log(FluTrain$ILI))**2))/length(FluTrain$ILI)

Rsquared1 = summary(FluTrend1)$r.squared
Rsquared1

#Manual calculation of Rsquarred
RSS.Trend1 <- sum(FluTrend1$residuals**2)
model0 = lm(log(ILI) ~ 1, data=FluTrain)
TSS <- sum(model0$residuals**2)
Rsquared.calc =  1-(RSS.Trend1 /TSS)
Rsquared.calc
```

## 2.3
```{r }
FluTest <- read.csv("FluTest.csv", header=FALSE)

colnames(FluTest) = c("Weeks", "ILI", "Queries")

head(FluTest)
```
### 2.3.6
```{r }
paste("Predicted value for 11th March 2012 : " ,exp(predict(FluTrend1, FluTest[FluTest$Weeks=="2012-03-11 - 2012-03-17",])))
```
### 2.3.7
```{r }
paste("The real value for 11th March 2012 : " ,FluTest[FluTest$Weeks=="2012-03-11 - 2012-03-17",]$ILI)
paste("Squared error : " ,(exp(predict(FluTrend1, FluTest[FluTest$Weeks=="2012-03-11 - 2012-03-17",]))-FluTest[FluTest$Weeks=="2012-03-11 - 2012-03-17",]$ILI)**2)

```
### 2.3.8
```{r }
sum(((exp(predict(FluTrend1, FluTest))-FluTest$ILI)**2))/length(FluTest$ILI)
```

## 2.4
### 2.4.9
```{r }
head(FluTrain)
FluTrain$ILI_Lag2 = c(c(NA, NA), as.vector(FluTrain$ILI[0:415]))
plot(log(FluTrain$ILI_Lag2)~FluTrain$ILI)
```

There is a strong positive relationship between log(ILI Lag2) and log(ILI).

### 2.4.10
```{r }
FluTrend2 = lm(log(ILI) ~ Queries + log(ILI_Lag2) , data = FluTrain)
summary(FluTrend2)

summary(FluTrend2)$r.squared
mseTrend2 = mean((predict(FluTrend2, FluTrain)-FluTrain$ILI)**2, na.rm=TRUE)
Rsquared2 = summary(FluTrend2)$r.squared
```
### 2.4.11
```{r }

mseTrend2<mseTrend1.1
Rsquared2>Rsquared1
```
Due to overfitting, FluTrend2 is a weaker model then FluTrend1 on the training set.

## 2.5

### 2.5.12

The ILI value of the second-to-last observation in the training data.

### 2.5.13
The ILI value of the last observation in the training data.

### 2.5.14

```{r }
FluTest$ILI_Lag2 = c( as.vector(FluTrain$ILI[416:417]), as.vector(FluTest$ILI[0:50]))
as.vector(FluTrain$ILI[416:417])
```

### 2.5.15
```{r }
test_predTrend2 = exp(predict(FluTrend2, FluTest))
mse.test2 = mean((test_predTrend2-FluTest$ILI)**2, na.rm=TRUE)
```
### 2.5.16
```{r }
test_predTrend1 = exp(predict(FluTrend1, FluTest))
mse.test1 = mean((test_predTrend1-FluTest$ILI)**2, na.rm=TRUE)
mse.test2<mse.test1
```
## 2.6
```{r }
arima(FluTest$ILI, order = c(3,0,0), method = "CSS")
```


